---
title: "asthma_map"
author: "Gunnar"
date: "12/8/2021"
output: github_document
---

```{r setup, include=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r}
library(tidyverse)
library(janitor)
library(haven)
library(plotly)
library(readr)
library(rjson)
```
##Creating state level brfss map of asthma prevalence
```{r}
#Loading the main brfss data from local library
state_df <- read_csv("brfss_data/brfss_epa.csv")
```

```{r}
#Have to use state abbreviations to work with the plotly map
state_df <- state_df %>%
  mutate(state = ifelse(state == 66, NA, state)) %>%
  drop_na() %>%
  mutate(state = factor(state, labels = c("AL", 'AK', "AZ", "AR", "CA", "CO", "CT", "DE", "DC", "FL", "GA", "HI", "ID", "IL", "IN", "IA", "KS", "KY", "LA", "ME", "MD", "MA", "MI", "MN", "MS", "MO", "MT", "NE", "NV", "NH", "NJ", "NM", "NY", "NC", "ND", "OH", "OK", "OR", "PA", "RI", "SC", 'SD', 'TN', 'TX', 'UT', 'VT', 'VA', 'WA', 'WV', 'WI', 'WY')))

#Creating an asthma prevalence variable
state_df <- state_df %>%
  group_by(state, iyear, median_aqi) %>%
  count(asthma_current) %>%
  mutate(asthma_perc = (n/sum(n) * 100), asthma_perc = round(asthma_perc, 2)) %>%
  filter(asthma_current == "yes")

#Creating hover label for map
state_df <- state_df %>%
  mutate(hover = paste0(state, " Asthma Prevalence: ", asthma_perc, "<br>", "Median AQI: ", median_aqi))
```

```{r}
#Creating the state level asthma prevalence map in plotly
geo_properties = list(
  scope = "usa",
  projection = list(
    type = "albers usa"),
  showland = TRUE,
  landcolor = toRGB('white'))

asthma_graph <- state_df %>%
  plot_geo(
    locationmode = 'USA-states',
    frame = ~iyear) %>%
  add_trace(locations = ~state,
            z = ~asthma_perc,
            zmin = 0,
            zmax = 15,
            text = ~hover,
            hoverinfo = 'text',
            color = ~asthma_perc,
            colorscale = "Viridis") %>%
  layout(geo = geo_properties,
         font = "DM Sans",
         title = "Asthma Prevalence in the United State, 2013-2021") %>%
  colorbar(title = "Asthma Prevalence") %>%
  animation_slider(
    currentvalue = list(prefix = "YEAR "))
```

##Creating a map of AQI by county by year in plotly
#First load and clean the epa dataset from local git

```{r}
#reuse state_df

#Creating the state level median aqi map in plotly
geo_properties = list(
  scope = "usa",
  projection = list(
    type = "albers usa"),
  showland = TRUE,
  landcolor = toRGB('white'))

aqi_graph <- state_df %>%
  plot_geo(
    locationmode = 'USA-states',
    frame = ~iyear) %>%
  add_trace(locations = ~state,
            z = ~median_aqi,
            zmin = 0,
            zmax = max(state_df$median_aqi),
            text = ~hover,
            hoverinfo = 'text',
            color = ~median_aqi,
            colorscale = "Viridis") %>%
  layout(geo = geo_properties,
         font = "DM Sans",
         title = "Median AQI in the United State, 2013-2021") %>%
  colorbar(title = "Median AQI") %>%
  animation_slider(
    currentvalue = list(prefix = "YEAR "))
```






##Creating a map of asthma prevalence by urban county by year in plotly
#First load the brfss_cities dataset from local brfss library
#Also load county level epa data
```{r}
cities_df <- read_csv("brfss_data/brfss_cities.csv")
#state abbreviations csv
state_ab <- read_csv("brfss_data/csvData.csv") %>%
  clean_names() %>%
  select(state, code)

epa_counties <- read_csv("brfss_data/epa.csv") %>%
  rename(state = state_name) %>%
  select(state, county, year, median_aqi)
  
epa_counties <- left_join(epa_counties, state_ab) %>%
  select(-state)

#join with county level data
county_data <- left_join(cities_df, epa_counties, by = c("county" = "county", "state" = "code", "brfss_year" = "year"))

county_data %>%
  filter(state == "CA")
```
#create a prevalence variable for asthma_current and then make a hover label
```{r}
county_data <- county_data %>%
  group_by(fips, brfss_year, county, state, median_aqi) %>%
  count(asthma_current) %>%
  mutate(asthma_perc = (n/sum(n) * 100), asthma_perc = round(asthma_perc, 2)) %>%
  filter(asthma_current == "yes") %>%
  ungroup()

county_data <- county_data %>%
  mutate(hover = paste0(county, ", ", state, " Asthma Prevalence: ", asthma_perc, "<br>", "Median AQI: ", median_aqi),
         nfips = sprintf("%05d", fips))
```

```{r}
#Load county level map from the below link
url <- 'https://raw.githubusercontent.com/plotly/datasets/master/geojson-counties-fips.json'

counties <- rjson::fromJSON(file=url)


#Creating the map
g <- list(
  scope = 'usa',
  projection = list(type = 'albers usa'),
  showlakes = TRUE,
  lakecolor = toRGB('white'))

urban_asthma_map <- county_data %>% 
  plot_ly(
    frame = ~brfss_year
  ) %>% 
  add_trace(
    type="choropleth",
    geojson=counties,
    locations = county_data$nfips,
    z = county_data$asthma_perc,
    colorscale="Viridis",
    zmin=0,
    zmax=15,
    text = ~hover,
    hoverinfo = "text",
    marker=list(line=list(
      width=0))) %>%
  layout(geo = g,
         font = "DM Sans",
         title = "Asthma Prevalence in US Urban Counties, 2013-2019") %>%
  colorbar(title = "Asthma Prevalence") %>%
    animation_slider(
    currentvalue = list(prefix = "YEAR "))
```

#creating urban county-level AQI map reusing urban_data df
```{r}
#Load county level map from the below link
url <- 'https://raw.githubusercontent.com/plotly/datasets/master/geojson-counties-fips.json'

counties <- rjson::fromJSON(file=url)


#Creating the map
g <- list(
  scope = 'usa',
  projection = list(type = 'albers usa'),
  showlakes = TRUE,
  lakecolor = toRGB('white'))

urban_aqi_map <- county_data %>% 
  plot_ly(
    frame = ~brfss_year
  ) %>% 
  add_trace(
    type="choropleth",
    geojson=counties,
    locations = county_data$nfips,
    z = county_data$median_aqi,
    colorscale="Viridis",
    zmin=0,
    zmax=max(county_data$median_aqi),
    text = ~hover,
    hoverinfo = "text",
    marker=list(line=list(
      width=0))) %>%
  layout(geo = g,
         font = "DM Sans",
         title = "Median AQI in US Urban Counties, 2013-2019") %>%
  colorbar(title = "Median AQI") %>%
    animation_slider(
    currentvalue = list(prefix = "YEAR "))
```


## State Level Maps of Asthma Prevalence and AQI
```{r}
aqi_graph

asthma_graph
```

## County Level Maps of Asthma Prevalence and AQI
```{r}
urban_aqi_map

urban_asthma_map
```



