---
title: "epa_cleaning_brfss_merge_2"
author: "Amanda Miles"
date: "12/8/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(dplyr)

```

## Reading in and Cleaning EPA Data

```{r epa_cleaning}

epa_df = 
  tibble(path = list.files("./data", full.names = TRUE)) %>% 
  mutate(data = purrr::map(.x = path, ~read_csv(.x))) %>% 
  janitor::clean_names() %>%
  unnest(data) %>%
  janitor::clean_names() %>%
  select(-c("path",
            "days_with_aqi",
            "max_aqi",
            "good_days":"hazardous_days", 
            "x90th_percentile_aqi")) %>%
  rename("days_pm2.5" = "days_pm2_5", "state_name" = "state") %>%
  mutate(state_code = state_name,
         state_code = str_replace_all(state_code, " ", ""),
         state_code = str_to_lower(state_code),
         state_code =  
           recode(state_code, 
             alabama = 1, alaska = 2, arizona = 4, arkansas = 5, 
             california = 6, colorado = 8, connecticut = 9, delaware = 10,
             districtofcolumbia = 11, florida = 12, georgia = 13, hawaii = 15, 
             idaho = 16, illinois = 17, indiana = 18, iowa = 19, kansas = 20, 
             kentucky = 21, louisiana = 22, maine = 23, maryland = 24, massachusetts = 25, 
             michigan = 26, minnesota = 27,mississippi = 28, missouri = 29, montana = 30, 
             nebraska = 31, nevada = 32, newhampshire = 33, newmexico = 35,
             newyork = 36, northcarolina = 37, northdakota = 38, ohio = 39, oklahoma = 40, 
             oregon = 41, pennsylvania = 42 , rhodeisland = 44, southcarolina = 45, 
             southdakota = 46, tennessee = 47, texas = 48, utah = 49, vermont = 50,
             virginia = 51, washington = 53, westvirginia = 54, wisconsin = 55, 
             wyoming = 56, puertorico = 72
             ),
         state_code = as.factor(state_code)) %>%
  na.omit(state_code) %>%
  relocate(state_code, state_name, everything())

# checking missing values
sapply(epa_df, function(x) sum(is.na(x)))

view(epa_df)

```

New Jersey, Guam, the Virgin Islands, and the City of Mexico do not have codes associated in this [codebook](https://www.cdc.gov/brfss/annual_data/2019/pdf/codebook19_llcp-v2-508.HTML). They have been removed from the EPA dataset.

## Creating an EPA State-Level AQI DF

```{r epa_state_level}

state_epa_df = epa_df %>%
  group_by(state_code, state_name, year) %>%
  summarize(median = median(median_aqi)) %>%
  rename("median_aqi" = "median") 

view(state_epa_df)

```

## Adding Column Combining State Code and Year for the Merge

```{r epa_state_year_code}

state_epa_df = state_epa_df %>%
  mutate(
    state_code_year = paste0(state_code, "_", year)
  ) %>%
  relocate(state_code_year, everything()) %>%
  select(-state_year)

head(state_epa_df)

```


## Merging EPA and BRFSS State-Level Data

* Read in the BRFSS data
* Made state a factor in the BRFSS DF
* Created a new column in BRFSS which concatenates state and year for the merge
* Merged EPA state-level data with BRFSS individual-level data

```{r epa_brfss_merge}

brfss_df = read_csv(file = "./brfss_data/brfss_12072021.csv")

brfss_factor_df = brfss_df %>%
  mutate(
    state = as.factor(as.character(state)),
    state_year = paste0(state, "_", iyear)
  ) %>%
  relocate(state_year, state, iyear, everything())

head(brfss_factor_df)

brfss_epa_df = 
  left_join(brfss_factor_df, state_epa_df, by = c("state_year" = "state_code_year"))

```

brfss_epa_df is the file that merged EPA state-level AQI data by year with BRFSS individual level data 

```{r}

head(brfss_factor_df)

```





