---
title: "scatterplots"
author: "Amanda Miles"
date: "12/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(dplyr)
library(plotly)
library(flexdashboard)
library(viridis)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d

```

# Code that Merges EPA & BRFSS

## Reading in and Cleaning EPA Data

```{r epa_cleaning}

epa_df = 
  tibble(path = list.files("./data", full.names = TRUE)) %>% 
  mutate(data = purrr::map(.x = path, ~read_csv(.x))) %>% 
  janitor::clean_names() %>%
  unnest(data) %>%
  janitor::clean_names() %>%
  select(-c("path",
            "days_with_aqi",
            "max_aqi",
            "good_days":"hazardous_days", 
            "x90th_percentile_aqi")) %>%
  rename("days_pm2.5" = "days_pm2_5", "state_name" = "state") %>%
  mutate(state_code = state_name,
         state_code = str_replace_all(state_code, " ", ""),
         state_code = str_to_lower(state_code),
         state_code =  
           recode(state_code, 
             alabama = 1, alaska = 2, arizona = 4, arkansas = 5, 
             california = 6, colorado = 8, connecticut = 9, delaware = 10,
             districtofcolumbia = 11, florida = 12, georgia = 13, hawaii = 15, 
             idaho = 16, illinois = 17, indiana = 18, iowa = 19, kansas = 20, 
             kentucky = 21, louisiana = 22, maine = 23, maryland = 24, massachusetts = 25, 
             michigan = 26, minnesota = 27,mississippi = 28, missouri = 29, montana = 30, 
             nebraska = 31, nevada = 32, newhampshire = 33, newmexico = 35,
             newyork = 36, northcarolina = 37, northdakota = 38, ohio = 39, oklahoma = 40, 
             oregon = 41, pennsylvania = 42 , rhodeisland = 44, southcarolina = 45, 
             southdakota = 46, tennessee = 47, texas = 48, utah = 49, vermont = 50,
             virginia = 51, washington = 53, westvirginia = 54, wisconsin = 55, 
             wyoming = 56, puertorico = 72
             ),
         state_code = as.factor(state_code)) %>%
  na.omit(state_code) %>%
  relocate(state_code, state_name, everything())

view(epa_df)

```

New Jersey, Guam, the Virgin Islands, and the City of Mexico do not have codes associated in this [codebook](https://www.cdc.gov/brfss/annual_data/2019/pdf/codebook19_llcp-v2-508.HTML). They have been removed from the EPA dataset.

## Creating an EPA State-Level AQI DF

```{r epa_state_level}

state_epa_df = epa_df %>%
  group_by(state_code, state_name, year) %>%
  summarize(median = median(median_aqi)) %>%
  rename("median_aqi" = "median") 

view(state_epa_df)

```

## Adding Column Combining State Code and Year for the Merge

```{r epa_state_year_code}

state_epa_df = state_epa_df %>%
  mutate(
    state_code_year = paste0(state_code, "_", year)
  ) %>%
  relocate(state_code_year, everything())

head(state_epa_df)

```

## Merging EPA and BRFSS State-Level Data

* Read in the BRFSS data
* Made state a factor in the BRFSS DF
* Created a new column in BRFSS which concatenates state and year for the merge
* Merged EPA state-level data with BRFSS individual-level data

```{r epa_brfss_merge}

brfss_df = read_csv(file = "./brfss_data/brfss_12072021.csv")

brfss_factor_df = brfss_df %>%
  mutate(
    state = as.factor(as.character(state)),
    state_year = paste0(state, "_", iyear)
  ) %>%
  relocate(state_year, state, iyear, everything())

head(brfss_factor_df)

brfss_epa_df = 
  left_join(brfss_factor_df, state_epa_df, by = c("state_year" = "state_code_year"))

brfss_epa_df = brfss_epa_df %>%
  rename(epa_year = year) %>%
  relocate(state_year, state, state_name, iyear, everything())

```

# Calculating Asthma Prevalence by State & Year

```{r asthma_prev}

brfss_epa_df = brfss_epa_df %>%
  mutate(state = ifelse(state == 66, NA, state)) %>%
  drop_na() %>%
  mutate(state = factor(state, labels = c("AL", 'AK', "AZ", "AR", "CA", "CO", "CT", "DE", "DC", "FL", "GA", "HI", "ID", "IL", "IN", "IA", "KS", "KY", "LA", "ME", "MD", "MA", "MI", "MN", "MS", "MO", "MT", "NE", "NV", "NH", "NJ", "NM", "NY", "NC", "ND", "OH", "OK", "OR", "PA", "RI", "SC", 'SD', 'TN', 'TX', 'UT', 'VT', 'VA', 'WA', 'WV', 'WI', 'WY')))

asthma_prev_df = brfss_epa_df %>%
  group_by(state_year, state, iyear, median_aqi) %>%
  count(asthma_current) %>%
  mutate(asthma_perc = (n/sum(n) * 100), asthma_perc = round(asthma_perc, 2)) %>%
  filter(asthma_current == "yes")
 
head(asthma_prev_df)

```

# Scatterplot of Asthma Prevalence and State Median AQI

```{r scatterplot}

asthma_prev_df %>%
  mutate(
    text_label = str_c("Year", iyear),
    state = fct_reorder(state, asthma_perc)) %>% 
  plot_ly(
  frame = ~iyear) %>%
  add_trace(
    type = "scatter",
    x = ~median_aqi,
    y = ~asthma_perc, 
    mode = "markers", 
    color = ~state, 
    colors = "viridis",
    text = ~text_label,
    alpha = 2
    ) %>%
  layout(
    xaxis = list(title = 'Median AQI'),
    yaxis = list(title = 'Asthma Prevalence (%)'),
    title = "Asthma Prevalence vs State Median AQI by Year"
  )
  
```

# Spearman Correlation Estimates: Asthma Prevalence and State Median AQI by Year

```{r spearman}

spearman_df = asthma_prev_df %>%
  relocate(iyear, everything())

spearman_nest = 
  nest(spearman_df, data = state_year:asthma_perc)

asthma_corr = function(df) {
  
  df %>%
  cor.test(
    median_aqi, 
    asthma_perc,
    method = "spearman")

}

spearman_nest %>% pull(data) %>% .[[1]]

spearman_nest = 
  spearman_nest %>% 
  mutate(
    correlation = map(.x = data, ~cor.test(.x$median_aqi, .x$asthma_perc, method = "spearman")),
    results = map(correlation, broom::tidy)) %>% 
  select(-data, -correlation) %>%
  unnest(results) %>%
  janitor::clean_names() %>%
  mutate(
    estimate = round(estimate, digits = 3),
    statistic = round(statistic, digits = 3),
    p_value = round(p_value, digits = 3)
  ) %>%
  select(-method, -alternative) %>%
  rename(year = iyear, test_statistic = statistic)

knitr::kable(spearman_nest, 
             digits = 2,
             caption = "*Spearman Correlation Estimates: Asthma Prevalence and State Median AQI by Year*")
  

```
# Spearman Correlation Estimates: Asthma Prevalence and State Median AQI from 2013-2020

```{r}


corr_results = cor.test(spearman_df$median_aqi, spearman_df$asthma_perc, method = "spearman")

corr_results = corr_results %>%
  broom::tidy() %>%
  janitor::clean_names() %>%
  mutate(
    estimate = round(estimate, digits = 3),
    statistic = round(statistic, digits = 3),
    p_value = round(p_value, digits = 3)
  ) %>%
  select(-method, -alternative) %>%
  rename(test_statistic = statistic)

view(corr_results)

knitr::kable(corr_results, 
             digits = 2,
             caption = "*Spearman Correlation Estimates: Asthma Prevalence and State Median AQI, 2013-2020*")


```
























































