---
title: <center>Summary Visualizations

output:   
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---
<br></br>

<center><a href=""><img src= "https://media.giphy.com/media/h542VAc7DgZdX62cLa/giphy.gif" width="240" />
</a></center>

<br></br>

# Asthma Prevalence and Air Quality Across the United States

```{r setup, include=FALSE, echo=FALSE, message=FALSE, warning=FALSE}

knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(tidyverse)
library(janitor)
library(haven)
library(plotly)
library(readr)
library(rjson)

##Creating state level brfss map of asthma prevalence

#Loading the main brfss data from local library
state_df <- read_csv("brfss_data/brfss_epa.csv")

#Have to use state abbreviations to work with the plotly map
state_df <- state_df %>%
  mutate(state = ifelse(state == 66, NA, state)) %>%
  drop_na() %>%
  mutate(state = factor(state, labels = c("AL", 'AK', "AZ", "AR", "CA", "CO", "CT", "DE", "DC", "FL", "GA", "HI", "ID", "IL", "IN", "IA", "KS", "KY", "LA", "ME", "MD", "MA", "MI", "MN", "MS", "MO", "MT", "NE", "NV", "NH", "NJ", "NM", "NY", "NC", "ND", "OH", "OK", "OR", "PA", "RI", "SC", 'SD', 'TN', 'TX', 'UT', 'VT', 'VA', 'WA', 'WV', 'WI', 'WY')))

#Creating an asthma prevalence variable
state_df <- state_df %>%
  group_by(state, iyear, median_aqi) %>%
  count(asthma_current) %>%
  mutate(asthma_perc = (n/sum(n) * 100), asthma_perc = round(asthma_perc, 2)) %>%
  filter(asthma_current == "yes")

#Creating hover label for map
state_df <- state_df %>%
  mutate(hover = paste0(state, " Asthma Prevalence: ", asthma_perc, "<br>", "Median AQI: ", median_aqi))

#Creating the state level asthma prevalence map in plotly
geo_properties = list(
  scope = "usa",
  projection = list(
    type = "albers usa"),
  showland = TRUE,
  landcolor = toRGB('white'))

asthma_graph <- state_df %>%
  plot_geo(
    locationmode = 'USA-states',
    frame = ~iyear) %>%
  add_trace(locations = ~state,
            z = ~asthma_perc,
            zmin = min(state_df$asthma_perc),
            zmax = 15,
            text = ~hover,
            hoverinfo = 'text',
            color = ~asthma_perc,
            colorscale = "Cividis") %>%
  layout(geo = geo_properties,
         font = "DM Sans",
         title = "Asthma Prevalence in the United State, 2013-2021") %>%
  colorbar(title = "Asthma Prevalence") %>%
  animation_slider(
    currentvalue = list(prefix = "YEAR "))

##Creating a map of AQI by county by year in plotly
#First load and clean the epa dataset from local git
#reuse state_df

#Creating the state level median aqi map in plotly
geo_properties = list(
  scope = "usa",
  projection = list(
    type = "albers usa"),
  showland = TRUE,
  landcolor = toRGB('white'))

aqi_graph <- state_df %>%
  plot_geo(
    locationmode = 'USA-states',
    frame = ~iyear) %>%
  add_trace(locations = ~state,
            z = ~median_aqi,
            zmin = min(state_df$median_aqi),
            zmax = max(state_df$median_aqi),
            text = ~hover,
            hoverinfo = 'text',
            color = ~median_aqi,
            colorscale = "Cividis") %>%
  layout(geo = geo_properties,
         font = "DM Sans",
         title = "Median AQI in the United State, 2013-2021") %>%
  colorbar(title = "Median AQI") %>%
  animation_slider(
    currentvalue = list(prefix = "YEAR "))

##Creating a map of asthma prevalence by urban county by year in plotly
#load the cleaned and combined county data + epa data
county_data <- read_csv("brfss_data/county_data.csv")

#Create a prevalence variable for asthma_current and then make a hover label

county_data <- county_data %>%
  group_by(fips, brfss_year, county, state, median_aqi) %>%
  count(asthma_current) %>%
  mutate(asthma_perc = (n/sum(n) * 100), asthma_perc = round(asthma_perc, 2)) %>%
  filter(asthma_current == "yes") %>%
  ungroup()

county_data <- county_data %>%
  mutate(hover = paste0(county, ", ", state, " Asthma Prevalence: ", asthma_perc, "<br>", "Median AQI: ", median_aqi),
         nfips = sprintf("%05d", fips))

#Load county level map from the below link
url <- 'https://raw.githubusercontent.com/plotly/datasets/master/geojson-counties-fips.json'

counties <- rjson::fromJSON(file=url)

#Creating the map
g <- list(
  scope = 'usa',
  projection = list(type = 'albers usa'),
  showlakes = TRUE,
  lakecolor = toRGB('white'))

urban_asthma_map <- county_data %>% 
  plot_ly(
    frame = ~brfss_year
  ) %>% 
  add_trace(
    type="choropleth",
    geojson=counties,
    locations = county_data$nfips,
    z = county_data$asthma_perc,
    colorscale="Cividis",
    zmin=min(county_data$asthma_perc),
    zmax=15,
    text = ~hover,
    hoverinfo = "text",
    marker=list(line=list(
      width=0))) %>%
  layout(geo = g,
         font = "DM Sans",
         title = "Asthma Prevalence in US Urban Counties, 2013-2019") %>%
  colorbar(title = "Asthma Prevalence") %>%
    animation_slider(
    currentvalue = list(prefix = "YEAR "))

#Creating urban county-level AQI map reusing urban_data df

#Load county level map from the below link
url <- 'https://raw.githubusercontent.com/plotly/datasets/master/geojson-counties-fips.json'

counties <- rjson::fromJSON(file=url)

#Creating the map
g <- list(
  scope = 'usa',
  projection = list(type = 'albers usa'),
  showlakes = TRUE,
  lakecolor = toRGB('white'))

urban_aqi_map <- county_data %>% 
  plot_ly(
    frame = ~brfss_year
  ) %>% 
  add_trace(
    type="choropleth",
    geojson=counties,
    locations = county_data$nfips,
    z = county_data$median_aqi,
    colorscale="Cividis",
    zmin=min(county_data$median_aqi),
    zmax=max(county_data$median_aqi),
    text = ~hover,
    hoverinfo = "text",
    marker=list(line=list(
      width=0))) %>%
  layout(geo = g,
         font = "DM Sans",
         title = "Median AQI in US Urban Counties, 2013-2019") %>%
  colorbar(title = "Median AQI") %>%
    animation_slider(
    currentvalue = list(prefix = "YEAR "))
```

### State Level Maps of Asthma Prevalence and AQI
```{r AQI and Asthma graphs}
aqi_graph

asthma_graph
```

These maps show state-level asthma prevalence and median AQI across the United States between 2013 and 2020. State-level median AQI appears to be the highest in the west coast, southwest, and northeast regions  through 2013 and 2020. Similarly, state-level asthma prevalence appears to be highest in the pacific northwest and the northeast regions between 2013 and 2020.

<br></br>

### County Level Maps of Asthma Prevalence and AQI

```{r}

urban_aqi_map

urban_asthma_map
```

These maps show metropolitan statistical area-level asthma prevalence and median AQI across the United States between 2013 and 2019. Counties represented on these maps reflect the asthma and AQI estimates for the MMA they belong to. These maps reveal that several areas have asthma and AQI levels much higher than the average of their state. For example, the median AQI in San Bernardino county, CA was 80 in 2020 while the state median was 43. Similarly, the asthma prevalence in Worcester county, MA in 2020 was 13.96% while the state prevalence was 10.83%.

<br></br>
<br></br>

# Asthma Prevalence vs. State Median AQI by Year
```{r prevalence vs state, include=FALSE, echo=FALSE, message=FALSE, warning=FALSE}

#Clean data
epa_df = 
  tibble(path = list.files("./data", full.names = TRUE)) %>% 
  mutate(data = purrr::map(.x = path, ~read_csv(.x))) %>% 
  janitor::clean_names() %>%
  unnest(data) %>%
  janitor::clean_names() %>%
  select(-c("path",
            "days_with_aqi",
            "max_aqi",
            "good_days":"hazardous_days", 
            "x90th_percentile_aqi")) %>%
  rename("days_pm2.5" = "days_pm2_5", "state_name" = "state") %>%
  mutate(state_code = state_name,
         state_code = str_replace_all(state_code, " ", ""),
         state_code = str_to_lower(state_code),
         state_code =  
           recode(state_code, 
             alabama = 1, alaska = 2, arizona = 4, arkansas = 5, 
             california = 6, colorado = 8, connecticut = 9, delaware = 10,
             districtofcolumbia = 11, florida = 12, georgia = 13, hawaii = 15, 
             idaho = 16, illinois = 17, indiana = 18, iowa = 19, kansas = 20, 
             kentucky = 21, louisiana = 22, maine = 23, maryland = 24, massachusetts = 25, 
             michigan = 26, minnesota = 27,mississippi = 28, missouri = 29, montana = 30, 
             nebraska = 31, nevada = 32, newhampshire = 33, newmexico = 35,
             newyork = 36, northcarolina = 37, northdakota = 38, ohio = 39, oklahoma = 40, 
             oregon = 41, pennsylvania = 42 , rhodeisland = 44, southcarolina = 45, 
             southdakota = 46, tennessee = 47, texas = 48, utah = 49, vermont = 50,
             virginia = 51, washington = 53, westvirginia = 54, wisconsin = 55, 
             wyoming = 56, puertorico = 72
             ),
         state_code = as.factor(state_code)) %>%
  na.omit(state_code) %>%
  relocate(state_code, state_name, everything())


## Creating an EPA State-Level AQI DF
state_epa_df = epa_df %>%
  group_by(state_code, state_name, year) %>%
  summarize(median = median(median_aqi)) %>%
  rename("median_aqi" = "median") 

## Adding Column Combining State Code and Year for the Merge

state_epa_df = state_epa_df %>%
  mutate(
    state_code_year = paste0(state_code, "_", year)
  ) %>%
  relocate(state_code_year, everything())

brfss_df = read_csv(file = "./brfss_data/brfss_12072021.csv")

brfss_factor_df = brfss_df %>%
  mutate(
    state = as.factor(as.character(state)),
    state_year = paste0(state, "_", iyear)
  ) %>%
  relocate(state_year, state, iyear, everything())

head(brfss_factor_df)

brfss_epa_df = 
  left_join(brfss_factor_df, state_epa_df, by = c("state_year" = "state_code_year"))

brfss_epa_df = brfss_epa_df %>%
  rename(epa_year = year) %>%
  relocate(state_year, state, state_name, iyear, everything())

# Calculating Asthma Prevalence by State & Year
brfss_epa_df = brfss_epa_df %>%
  mutate(state = ifelse(state == 66, NA, state)) %>%
  drop_na() %>%
  mutate(state = factor(state, labels = c("AL", 'AK', "AZ", "AR", "CA", "CO", "CT", "DE", "DC", "FL", "GA", "HI", "ID", "IL", "IN", "IA", "KS", "KY", "LA", "ME", "MD", "MA", "MI", "MN", "MS", "MO", "MT", "NE", "NV", "NH", "NJ", "NM", "NY", "NC", "ND", "OH", "OK", "OR", "PA", "RI", "SC", 'SD', 'TN', 'TX', 'UT', 'VT', 'VA', 'WA', 'WV', 'WI', 'WY')))

asthma_prev_df = brfss_epa_df %>%
  group_by(state_year, state, iyear, median_aqi) %>%
  count(asthma_current) %>%
  mutate(asthma_perc = (n/sum(n) * 100), asthma_perc = round(asthma_perc, 2)) %>%
  filter(asthma_current == "yes")
```


```{r scatterplot}
# Scatterplot of Asthma Prevalence and State Median AQI

asthma_prev_df %>%
  mutate(
    text_label = str_c("Year", iyear),
    state = fct_reorder(state, asthma_perc)) %>% 
  plot_ly(
  frame = ~iyear) %>%
  add_trace(
    type = "scatter",
    x = ~median_aqi,
    y = ~asthma_perc, 
    mode = "markers", 
    color = ~state, 
    colors = "viridis",
    text = ~text_label,
    alpha = 2
    ) %>%
  layout(
    xaxis = list(title = 'Median AQI'),
    yaxis = list(title = 'Asthma Prevalence (%)'),
    title = "Asthma Prevalence vs State Median AQI by Year"
  )
```

* This scatterplot showed the relationship between median AQI and asthma prevalence by state and year. 
* For each year from 2013 to 2020, the data points were fairly dispersed and did not exhibit a monotonic relationship. There were also several outliers observed for each of the years examined. 
* These results differ from what was expected, since the team expected to see a monotonic relationship between median AQI and asthma prevalence. This may be due to the limitations we have in our data. We are primarily limited to analyzing asthma prevalence at the state level since BRFSS does not disclose asthma prevalence for non-metropolitan counties. Looking at the relationship grouped at the state level may mask the relationship between median AQI and asthma prevalence at the sub-state or county level. Additionally, the team’s ability to detect a relationship between the two variables may also be due to our data only containing information on annual median AQI rather than daily AQI. 


## Median AQI Over Time

```{r, echo = FALSE, message = FALSE, warning = FALSE}

knitr::opts_chunk$set(echo = FALSE)

epa_df = 
  tibble(path = list.files("./data", full.names = TRUE)) %>% 
  mutate(data = purrr::map(.x = path, ~read_csv(.x))) %>% 
  janitor::clean_names() %>%
  unnest(data) %>%
  janitor::clean_names() %>%
  select(-c("path",
            "days_with_aqi",
            "max_aqi",
            "good_days":"hazardous_days", 
            "x90th_percentile_aqi")) %>%
  rename("days_pm2.5" = "days_pm2_5") %>%
  mutate(state_name = state,
         state_name = str_replace_all(state_name, " ", ""),
         state_name = str_to_lower(state_name),
         state_name =  
           recode(state_name, 
             alabama = 1, alaska = 2, arizona = 3, arkansas = 5, 
             california = 6, colorado = 8, connecticut = 9, delaware = 10,
             districtofcolumbia = 11, florida = 12, georgia = 13, hawaii = 15, 
             idaho = 16, illinois = 17, indiana = 18, iowa = 19, kansas = 20, 
             kentucky = 21, louisiana = 22, maine = 23, maryland = 24, massachusetts = 25, 
             michigan = 26, minnesota = 27,mississippi = 28, missouri = 29, montana = 30, 
             nebraska = 31, nevada = 32, newhampshire = 33, newmexico = 35,
             newyork = 36, northcarolina = 37, northdakota = 38, ohio = 39, oklahoma = 40, 
             oregon = 41, pennsylvania = 42 , rhodeisland = 44, southcarolina = 45, 
             southdakota = 46, tennessee = 47, texas = 48, utah = 49, vermont = 50,
             virginia = 51, washington = 53, westvirginia = 54, wisconsin = 55, 
             wyoming = 56, puertorico = 72
             ),
         state_name = as.factor(state_name)) %>%
  na.omit(state_name) %>%
  relocate(state_name, state, everything())

state_aqi_df = epa_df %>%
  group_by(state_name, state, year) %>%
  summarize(median = median(median_aqi)) %>%
  rename("median_aqi" = "median") 
```

```{r EPA V2 Plot}
#Utilizing cleaned EPA data (v2) to create plots showing all states and median AQI over time:

state_aqi_df %>%
  group_by(state) %>%
  plot_ly() %>% 
  add_trace(type = 'scatter', 
            mode = 'lines+markers',
            x = ~year, 
            y = ~median_aqi, 
            color = ~state, 
            colors = 'viridis') %>%
  layout(xaxis = list(title = "Year"),
         yaxis = list(title = "Median AQI"),
         title = "Change in median AQI over time by state (2013 - 2020)")
```

```{r subplot code}

#Preparing for subplot by region

region_df = 
  state_aqi_df %>%
  mutate(region = state,
         region = as.character(region),
         #New England
         region = ifelse(region %in% c("Connecticut", "Maine",
                           "Massachusetts", "New Hampshire",
                           "Rhode Island", "Vermont"), 
                         "New England", region),
         #Middle Atlantic
         region = ifelse(region %in% c("New York", "Pennsylvania"), 
                         "Middle Atlantic", region),
         #East North Central
          region = ifelse(region %in% c("Indiana", "Illinois", "Michigan",
                           "Ohio", "Wisconsin"), 
                         "East North Central", region),
         #West North Central
         region = ifelse(region %in% c("Iowa", "Kansas","Minnesota", "Missouri",
                                       "Nebraska", "North Dakota", "South Dakota"), 
                         "West North Central", region),
         #South Atlantic
         region = ifelse(region %in%  c("Delaware", "District Of Columbia", "Florida", 
                                        "Georgia", "Maryland","North Carolina", "South Carolina", 
                                        "Virginia", "West Virginia", "Puerto Rico"), 
                         "South Atlantic & Puerto Rico", region),
         #East South Central
         region = ifelse(region %in% c("Alabama", "Kentucky","Mississippi", "Tennessee"), 
                         "East South Central", region),
         #West South Central
         region = ifelse(region %in% c("Arkansas", "Louisiana", "Oklahoma", "Texas"), 
                         "West South Central", region),
         #Mountain
         region = ifelse(region %in% c("Arizona", "Colorado", "Idaho", "New Mexico",
                                       "Montana", "Utah", "Nevada", "Wyoming"),
                         "Mountain", region),
         #Pacific
         region = ifelse(region %in% c("Alaska", "California", "Hawaii", "Oregon", "Washington"),
                         "Pacific", region)
         )
  
# To check missing values:
# sapply(region_df, function(x) sum(is.na(x)))

write.csv(region_df, "region.csv", row.names = TRUE)

#Plot for New England only

p1 = 
  region_df %>%
  group_by(state) %>%
  filter(region == "New England") %>%
  plot_ly() %>% 
  add_trace(type = 'scatter', 
            mode = 'lines+markers',
            x = ~year, 
            y = ~median_aqi, 
            color = ~state, 
            colors = 'viridis') %>%
  layout(xaxis = list(title = "Year"),
         yaxis = list(title = "Median AQI"))

#Plot for Middle Atlantic only

p2 =
  region_df %>%
  group_by(state) %>%
  filter(region == "Middle Atlantic") %>%
  plot_ly() %>% 
  add_trace(type = 'scatter', 
            mode = 'lines+markers',
            x = ~year, 
            y = ~median_aqi, 
            color = ~state, 
            colors = 'viridis') %>%
  layout(xaxis = list(title = "Year"),
         yaxis = list(title = "Median AQI"))


#Plot for East North Central only

p3 =
  region_df %>%
  group_by(state) %>%
  filter(region == "East North Central") %>%
  plot_ly() %>% 
  add_trace(type = 'scatter', 
            mode = 'lines+markers',
            x = ~year, 
            y = ~median_aqi, 
            color = ~state, 
            colors = 'viridis') %>%
  layout(xaxis = list(title = "Year"),
         yaxis = list(title = "Median AQI"))


#Plot for West North Central only

p4 = 
  region_df %>%
  group_by(state) %>%
  filter(region == "West North Central") %>%
  plot_ly() %>% 
  add_trace(type = 'scatter', 
            mode = 'lines+markers',
            x = ~year, 
            y = ~median_aqi, 
            color = ~state, 
            colors = 'viridis') %>%
  layout(xaxis = list(title = "Year"),
         yaxis = list(title = "Median AQI"))

#Plot for South Atlantic & Puerto Rico

p5 = 
  region_df %>%
  group_by(state) %>%
  filter(region == "South Atlantic & Puerto Rico") %>%
  plot_ly() %>% 
  add_trace(type = 'scatter', 
            mode = 'lines+markers',
            x = ~year, 
            y = ~median_aqi, 
            color = ~state, 
            colors = 'viridis') %>%
  layout(xaxis = list(title = "Year"),
         yaxis = list(title = "Median AQI"))

#Plot for East South Central only

p6 = 
  region_df %>%
  group_by(state) %>%
  filter(region == "East South Central") %>%
  plot_ly() %>% 
  add_trace(type = 'scatter', 
            mode = 'lines+markers',
            x = ~year, 
            y = ~median_aqi, 
            color = ~state, 
            colors = 'viridis') %>%
  layout(xaxis = list(title = "Year"),
         yaxis = list(title = "Median AQI"))

#Plot for West South Central only

p7 = 
  region_df %>%
  group_by(state) %>%
  filter(region == "West South Central") %>%
  plot_ly() %>% 
  add_trace(type = 'scatter', 
            mode = 'lines+markers',
            x = ~year, 
            y = ~median_aqi, 
            color = ~state, 
            colors = 'viridis') %>%
  layout(xaxis = list(title = "Year"),
         yaxis = list(title = "Median AQI"))

#Plot for Mountain only

p8 =
  region_df %>%
  group_by(state) %>%
  filter(region == "Mountain") %>%
  plot_ly() %>% 
  add_trace(type = 'scatter', 
            mode = 'lines+markers',
            x = ~year, 
            y = ~median_aqi, 
            color = ~state, 
            colors = 'viridis') %>%
  layout(xaxis = list(title = "Year"),
         yaxis = list(title = "Median AQI"))

#Plot for Pacific only

p9 = 
  region_df %>%
  group_by(state) %>%
  filter(region == "Pacific") %>%
  plot_ly() %>% 
  add_trace(type = 'scatter', 
            mode = 'lines+markers',
            x = ~year, 
            y = ~median_aqi, 
            color = ~state, 
            colors = 'viridis') %>%
  layout(xaxis = list(title = "Year"),
         yaxis = list(title = "Median AQI"))

#Another subplot method. However, using this method we are not able to provide titles for each subplot as we can in ggplot2 facet grid

#region_df %>%
#  group_by(region) %>%
#  do(p = plot_ly( ., 
#            x = ~year, 
#            y = ~median_aqi,
#            color = ~state, 
#            #colors = 'viridis',
#            type = 'scatter', 
#            mode = 'lines+markers')) %>%
#  subplot(nrows = 4, shareX = FALSE, shareY = TRUE) %>%
#  layout(xaxis = list(title = "Year"),
#         yaxis = list(type = "Median AQI"),
#         title = "Change in median AQI over time by region (2013 - 2020)")

#Putting together plots by region & formatting subplot titles

plot = 
  subplot(p1, p2, p3, p4, p5, p6, p7, p8, p9,
        nrows = 3,
        shareY = T,
        shareX = T) %>%
  layout(title = "Median AQI by U.S. state region (2013 - 2020)")
        
annotations = list( 
  list( 
    x = 0.15,  
    y = 0.99,  
    text = "New England",  
    xref = "paper",  
    yref = "paper",  
    xanchor = "center",  
    yanchor = "bottom",  
    showarrow = FALSE 
  ),  
  list( 
    x = 0.50,  
    y = 0.99,  
    text = "Middle Atlantic",  
    xref = "paper",  
    yref = "paper",  
    xanchor = "center",  
    yanchor = "bottom",  
    showarrow = FALSE 
  ),  
  list( 
    x = 0.85,  
    y = 0.99,  
    text = "East North Central",  
    xref = "paper",  
    yref = "paper",  
    xanchor = "center",  
    yanchor = "bottom",  
    showarrow = FALSE 
  ),
  list( 
    x = 0.15,  
    y = 0.63,  
    text = "West North Central",  
    xref = "paper",  
    yref = "paper",  
    xanchor = "center",  
    yanchor = "bottom",  
    showarrow = FALSE 
  ),
  list( 
    x = 0.50,  
    y = 0.63,  
    text = "South Atlantic & Puerto Rico",  
    xref = "paper",  
    yref = "paper",  
    xanchor = "center",  
    yanchor = "bottom",  
    showarrow = FALSE 
  ),
  list( 
    x = 0.85,  
    y = 0.63,  
    text = "East South Central",  
    xref = "paper",  
    yref = "paper",  
    xanchor = "center",  
    yanchor = "bottom",  
    showarrow = FALSE 
  ),
  list( 
    x = 0.15,  
    y = 0.30,  
    text = "West South Central",  
    xref = "paper",  
    yref = "paper",  
    xanchor = "center",  
    yanchor = "bottom",  
    showarrow = FALSE 
  ),
  list( 
    x = 0.50,  
    y = 0.30,  
    text = "Mountain",  
    xref = "paper",  
    yref = "paper",  
    xanchor = "center",  
    yanchor = "bottom",  
    showarrow = FALSE 
  ),
  list( 
    x = 0.85,  
    y = 0.30,  
    text = "West South Central",  
    xref = "paper",  
    yref = "paper",  
    xanchor = "center",  
    yanchor = "bottom",  
    showarrow = FALSE 
  ))

plot <- plot %>% 
  layout(annotations = annotations) 
```

```{r plot region}
plot
```

* The interactive spaghetti plots depict the trend in median AQI over time by each state, and further stratified by [regions of the United States](https://www2.census.gov/geo/pdfs/maps-data/maps/reference/us_regdiv.pdf).  The trends in median AQI over time per region in the United States decreased from 2013 to 2020, with the exception of Puerto Rico. The state that had the largest decrease in median AQI was Pennsylvania, with a median AQI of 44 in 2013 and a median AQI of 37 in 2020. The state with the highest median AQI over time was California, with relatively little change in median AQI over time. 


